---
title: "Homework Assignment #2"
subtitle: "Part 2"
author: "Liane Chen"
date: 2024-02-03
toc: true
format: html
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                    setup                                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#..........................load packages.........................
library(tidyverse)
library(sf)
library(dplyr)
library(lubridate)
library(janitor)
library(ggplot2)
library(naniar)

#..........................import data...........................
wildfire_original <- st_read("data/mtbs_perims_DD.shp")
```

```{r}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                wrangle data                              ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

wildfire <- janitor::clean_names(wildfire_original) %>% 
  
  #replace the numeric '-99999' code with NA values in order to drop NAs
  replace_with_na(replace = list(-99999)) %>% 
  
  #select columns we are interested in
  select(incid_name, incid_type, burn_bnd_ac, ig_date, geometry) %>% 
  
  #remove NA observations
  drop_na()

```


# Let's do some exploratory visuals! ----

```{r}
#| eval: true
#| echo: true

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                  plot #1                                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Try a histogram with facet wrap
wildfire %>% 
  #mutate(month_name = factor(month_name, levels = month.name)) %>%  #organize month names
  ggplot(aes(x = ig_date)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 1) + #by default info is stacked, so let's add position by identity and make it transparent so it's clear
  facet_wrap(~incid_type) +
  #adjust bins so you can see better
  scale_fill_manual(values = c("#2C5374", "#ADD8E6", "#8B3A3A"))

```

```{r}

#| eval: true
#| echo: true

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                  plot #2                                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# try bar plot

wildfire %>% 
  ggplot(aes(x = burn_bnd_ac, y = incid_type)) + 
  geom_col(fill = "purple") + 
  coord_flip()
```


```{r}
#| eval: true
#| echo: true

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                  plot #3                                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#try a violin

wildfire %>% 
  ggplot(aes(x=incid_type, y=burn_bnd_ac, fill=burn_bnd_ac)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.4, alpha=0.9) 

```
```{r}
#| eval: true
#| echo: true

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                  plot #4                                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# try histogram

wildfire %>%
  ggplot( aes(x=ig_date)) +
    geom_histogram() 
```


1. *What have you learned about your data? Have any potentially interesting patterns emerged? (5-8 sentences)*
```
From the initial data exploration, I learned that prescribed burns is considered 
a type of wildfire. I did not think of these as being in the same category 
before. The same can be said for wildland fire use burn types, which I would 
like to look into more. There are a good amount of unknown fire types that I 
wish were less so. Lastly, there is a clear trend of increased wildfire events 
in the last twenty years, which really sucks, in scientific terms. The histogram 
shows this pretty clearly. 

```

2. *In HW #1, you outlined some questions that you wanted to answer using these data. Have you made any strides towards answering those questions? If yes, how so? If no, what next steps do you need to take (e.g. I need to create X plot type, I still need to track down Y data, I need to restructure existing data so that you can visualize it in Z ways, etc.)? (8-12 sentences)*
```
In assignment 1, I was hopeful that I could compare this dataset against US 
policy data to see if the creation of wildfire policies have had an impact on 
the actual occurances of wildfires in CA. However, I realized that the data for 
polcies was not extensive enough to be actually helpful. There were only like 2 
observations or so in that dataset. I decided to keep things contained and use 
this wildfire dataset only. It is possible that I can continue tracking down 
data. However, the reality may be that wildfire polciies are not that plentiful, 
so it may be hard to analyze in the way I want to here. I think that these 
initial data explorations have provided an initial sense of when burns are 
increasing. So perhaps I will rephrase my question. I think by looking at the 
geospatial information next, I should be able to find more interesting trends. 
I am guessing there will be trends towards the areas where I suspect fires have 
happened, where there are increased wildland and human interfaces. 

```

3. *What challenges do you foresee encountering with your data? These can be data wrangling and / or visualization challenges. (4-6 sentences)*
```
I foresee the challenge being able to relay the wildfire incident types with 
something useful, like location or dates. The dates starting from 1985 does 
seem to be a potential challenge as well. I think I will play around with the 
locations a little more too to see if I can gather more trends from that 
visualization. There are a lot of values too, which makes creating the plot 
harder too. I look forward to trying. 

```

