---
title: "Homework Assignment #3"
author: "Liane Chen"
date: 2024-02-03
toc: true
format:
  html:
    embed-resources: true
---

```{r setup, include=FALSE}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                    setup                                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#..........................load packages.........................
library(tidyverse)
library(sf)
library(dplyr)
library(lubridate)
library(janitor)
library(ggplot2)
library(naniar)
library(maps)
library(leaflet)

#..........................import data...........................
wildfire_original <- st_read("data/mtbs_perims_DD.shp")
```

```{r}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                wrangle data                              ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

wildfire <- janitor::clean_names(wildfire_original) %>% 
  
  #replace the numeric '-99999' code with NA values in order to drop NAs
  replace_with_na(replace = list(-99999)) %>% 
  
  #select columns we are interested in
  select(incid_name, incid_type, burn_bnd_ac, ig_date, geometry) %>% 
  
  #remove NA observations
  drop_na()

```

Which option do you plan to pursue? Restate your question(s).

Has this changed at all since HW #1? If yes, how so?

Explain which variables from your data set(s) you will use to answer your question(s).

Find at least two data visualizations that you could (potentially) borrow / adapt pieces from. Link to them or download and embed them into your .qmd file, and explain which elements you might borrow (e.g. the graphic form, legend design, layout, etc.).

Hand-draw your anticipated three visualizations (option 1) or infographic (option 2). Take a photo of your drawing and embed it in your rendered .qmd file – note that these are not exploratory visualizations, but rather your plan for your final visualizations that you will eventually polish and submit with HW #4.

Mock up your visualizations using code. 


What challenges did you encounter or anticipate encountering as you continue to build / iterate on your visualizations in R?

What ggplot extension tools / packages do you need to use to build your visualizations? Are there any that we haven’t covered in class that you’ll be learning how to use for your visualizations?

What feedback do you need from the instructional team and / or your peers to ensure that your intended message is clear?
