---
title: "Homework Assignment #3"
author: "Liane Chen"
date: 2024-02-03
toc: true
format:
  html:
    embed-resources: true
---

```{r setup, include=FALSE}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                    setup                                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#..........................load packages.........................
library(tidyverse)
library(sf)
library(dplyr)
library(lubridate)
library(janitor)
library(ggplot2)
library(naniar)
library(maps)
library(leaflet)
library(lubridate)
library(scales)
library(jsonlite)
library(gghighlight)

#..........................import data...........................
wildfire_original <- st_read("data/mtbs_perims_DD.shp")
```

```{r}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                wrangle data                              ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

wildfire <- janitor::clean_names(wildfire_original) %>% 
  
  #replace the numeric '-99999' code with NA values in order to drop NAs
  replace_with_na(replace = list(-99999)) %>% 
  
  #select columns we are interested in
  select(incid_name, incid_type, burn_bnd_ac, ig_date, geometry) %>% 
  
  #remove NA observations
  drop_na()

```

##Which option do you plan to pursue? Restate your question(s). Has this changed at all since HW #1? If yes, how so?

```         
ANSWER: I plan to puruse option 1, where I will have one question that each of the three visualizations works to answer, where each visualization is meant to reach / serve a different target audience and purpose. To restate, my question is: "How have wildfires changed over time?" Yes, this question has changed, since my first question in HW#1 relied on incorporating a dataset that turned out to be a bust and had very little information in it to extract anything meaningful. 
```

##Explain which variables from your data set(s) you will use to answer your question(s).

```         
ANSWER: The variables I will use to answer my question include 'ignition date', 'burned area (acres)', and 'incident types'.
```

##Find at least two data visualizations that you could (potentially) borrow / adapt pieces from. Link to them or download and embed them into your .qmd file, and explain which elements you might borrow (e.g. the graphic form, legend design, layout, etc.).

```
ANSWER: I like how Meghan annotates her plot with further context, while keeping the overall aethetic pretty clean in her hockey plot: https://meghan.rbind.io/blog/2022-02-22-increasing-the-flexibility-and-robustness-of-plots-in-ggplot2/ .I may borrow how she annotates her plot in the lower right corner of the plot. 

Another plot that I like is Cedric's historical mobile subscriptions plot here: https://raw.githubusercontent.com/Z3tt/TidyTuesday/main/plots/2020_46/2020_46_HistoricalPhones.png. I am thinking that perhaps this will be a good crossover for the histortical wildfire plots that I am trying to make. Specifically, I like his layout with the world proportions on the right, which I think will work nicely for the incident types, perhaps?

```

##Hand-draw your anticipated three visualizations (option 1) or infographic (option 2). Take a photo of your drawing and embed it in your rendered .qmd file -- note that these are not exploratory visualizations, but rather your plan for your final visualizations that you will eventually polish and submit with HW #4.

##Mock up your visualizations using code.

```{r}
wildfire <- unique(wildfire)

large_wildfire <- filter(wildfire, burn_bnd_ac >= 1000)

summary(large_wildfire)

large_wildfire <- na.omit(large_wildfire)

summary(large_wildfire)

# grouped_fire <- group_by(large_wildfire, incid_type, ig_date)
# 
# mean_burn_ac <- mean(grouped_fire$burn_bnd_ac, na.rm = TRUE)
# 
# summary(large_wildfire$burn_bnd_ac)
#                        
# 
# ggplot(data=grouped_fire, mapping = aes(x=ig_date, y=incid_type)) + geom_point() +
# geom_smooth(method=lm, se=TRUE) 


#+ ggtitle(“Average Size of Wildfires Has
```


###Visualization for a general audience
In this plot, we will include more data points from overall wildfire dataset, as to trigger more reaction when 
they see how counts have increased so much in later years. 
```{r}
large_wildfire %>%
  ggplot( aes(x=ig_date)) +
    geom_histogram(binwidth = 7, fill = "darkred", color = "darkred", alpha = 0.5) +
  labs(x = "Ignition Date", y = "Count", title = "Wildfire Incident Counts between 1984 and 2023") +
  theme_ipsum()
```




###Visualization for inclusion in a presentation
We will group the date into decades, so to make the date information easier to 
digest within a short amount of time. 
```{r}
#by decade

wildfire$year <- year(wildfire$ig_date)

wildfire_decade <- mutate(wildfire, 
                          decade = case_when(
                          year %in% 1980:1989 ~ "1980-1989",
                          year %in% 2000:2009 ~ "2000-2009", 
                          year %in% 1990:1999 ~ "1990-1999",
                          year %in% 2000:2009 ~ "2000-2009", 
                          year %in% 2010:2019 ~ "2010-2019",
                          year %in% 2020:2029 ~ "2020-2023",
                                 TRUE ~ NA_character_
                          ))

wildfire_decade <- group_by(wildfire_decade, decade, na.rm=TRUE)

ggplot(data=wildfire_decade, width = 4, height = 2) + 
  geom_col(mapping = aes(x=decade, y=burn_bnd_ac), fill="darkred", width=0.8) +
  scale_y_continuous(name ="Acreages Burned (acres)", labels = scales::comma) +
  xlab("Decade") +
  annotate("text", x=Inf, y=Inf, hjust=1, vjust=4.5,
           label = "*Note that data\nis available only\nthrough year 2023", family = "Avenir") +
  ggtitle("Wildfire Acreages Burned by Decade") +
  theme(plot.title = element_text(size = 16, face = "bold", margin = margin(b = 14)),
        panel.background = element_rect(fill = "white"))  
        
```


```{r}
library(hrbrthemes)

# Plot
ggplot(largest_wildfire, aes(x=xValue, y=yValue)) +
  geom_area( fill="#69b3a2", alpha=0.4) +
  geom_line(color="#69b3a2", size=2) +
  geom_point(size=3, color="#69b3a2") +
  theme_ipsum() +
  ggtitle("Evolution of something")
```

```{r}
library(ggstream)

# Stacked area chart with smoothing
largest_wildfire %>% 
  ggplot(aes(year, burn_bnd_ac, fill = burn_bnd_ac, label=incid_name)) +
  geom_area()
  #+ #color = incid_type)) +
  #geom_stream(streamtype = "ridge", aes(bw= 1)) 


```


```{r}
largest_wildfire %>%
  ggplot(aes(x = year, y = burn_bnd_ac, fill = as.factor(burn_bnd_ac), label = incid_name)) +
  geom_area(position = "stack") #+
  geom_text(aes(y = 0, label = incid_name), vjust = 0, hjust = 1, size = 2, color = "black") +
  scale_fill_continuous(guide = "none") +  # You can customize the color scale
  labs(title = "Wildfire Data",
       x = "Year",
       y = "Burned Area (acres)",
       fill = "Incident Name") +
  theme_minimal()
```



##What challenges did you encounter or anticipate encountering as you continue to build / iterate on your visualizations in R?

##What ggplot extension tools / packages do you need to use to build your visualizations? Are there any that we haven't covered in class that you'll be learning how to use for your visualizations?

##What feedback do you need from the instructional team and / or your peers to ensure that your intended message is clear?

```{r}
# load United States state map data
MainStates <- map_data("state")

CA <- MainStates %>% 
  filter(region == "california")

largest_wildfire <- large_wildfire %>% 
  filter(incid_type == "Wildfire") %>% 
  arrange(desc(burn_bnd_ac)) %>% 
  slice(1:20)

# add year column to data frame
largest_wildfire$year <- format(largest_wildfire$ig_date, "%Y")

largest_wildfire <- transform(largest_wildfire, year = as.numeric(year))

largest_wildfire %>%
  ggplot(aes(x = year, y = burn_bnd_ac, fill = burn_bnd_ac)) +
  geom_area(position = "stack", fill="darkred") + 
  scale_fill_continuous(guide = "none") +  
  labs(title = "Wildfire Incidents over the Years",
       x = "Year",
       y = "Burned Area (acres)",
       fill = "Incident Name") +
  theme_minimal()




```
```{r}
large_wildfire %>%
  ggplot(aes(x = year, y = burn_bnd_ac, fill = burn_bnd_ac)) +
  geom_area(position = "stack", fill="darkred") + 
  scale_fill_continuous(guide = "none") +  
  labs(title = "Wildfire Incidents over the Years",
       x = "Year",
       y = "Burned Area (acres)",
       fill = "Incident Name") +
  theme_minimal()
```

```{r}

# add year column to data frame
large_wildfire$year <- format(large_wildfire$ig_date, "%Y")

large_wildfire <- transform(large_wildfire, year = as.numeric(year))

large_wildfire %>%
  ggplot(aes(x = year, y = burn_bnd_ac, fill = burn_bnd_ac)) +
  #geom_area(position = "stack", fill="darkred") + 
  geom_smooth(color="darkred") +
  scale_fill_continuous(guide = "none") +  
  labs(title = "Wildfire Incidents between 1984 and 2023",
       x = "Year",
       y = "Burned Area (acres)") +
  theme_minimal()
```



```{r}
US_wildfires <- largest_wildfire %>% 
  filter(!(incid_name %in% 
           c("RAILBELT COMPLEX (MINTO FLATS SOUTH)", "CHALKYITSIK COMPLEX", "TAYLOR HIGHWAY COMPLEX (BILLY CREEK)")))
```

```{r}
ggplot() +
  geom_polygon(data=CA, aes(x=long, y=lat, group=group),
                color="black", fill="lightgrey" )+
  geom_sf(data = US_wildfires, aes(fill = burn_bnd_ac), color="darkred")

##can look up which ones are in alaska and delete from dataframe
## alterntatively, look up which ones are in CA

```
